### 2022-05-17 masters_paper_new_2022(1).tex
Add files via upload

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\author{}
\date{}

\begin{document}

\begin{quote}
Predicting the Quality of CDRH3 Antibody Loop Structural Models

From the Department of Structural and Molecular Biology,

University College of London, Gower Street, London WC1E 6BT UK
\end{quote}

\hypertarget{abstract}{%
\section{ABSTRACT}\label{abstract}}

\hypertarget{therapeutic-antibodies-have-shown-an-unprecedented-pace-of-development-and-have-brought-new-hope-for-the-treatment-of-numerous-diseases.-the-bioinformatic-tools-for-modelling-antibody-structures-have-become-invaluable-for-antibody-engineering-and-the-development-of-therapeutic-antibodies.-the-antigen-binding-site-consists-of-six-hypervariable-loops-also-known-as-the-complementary-determining-regions-cdr-all-of-which-can-be-modelled-with-adequate-accuracy-except-for-one.-it-remains-markedly-difficult-to-model-the-third-cdr-loop-of-the-antibody-heavy-chain.-the-cdrh3-differs-in-length-has-far-greater-sequence-variability-and-has-such-a-great-structural-diversity-that-modelling-it-is-considerably-harder.}{%
\section{\texorpdfstring{Therapeutic antibodies have shown an
unprecedented pace of development and have brought new hope for the
treatment of numerous diseases. The bioinformatic tools for modelling
antibody structures have become invaluable for antibody engineering and
the development of therapeutic antibodies. The antigen-binding site
consists of six hypervariable loops, also known as the Complementary
Determining Regions (CDR), all of which can be modelled with adequate
accuracy, except for one. It remains markedly difficult to model the
third CDR loop of the antibody heavy chain. The CDRH3 differs in length,
has far greater sequence variability and has such a great structural
diversity that modelling it is considerably harder.
}{Therapeutic antibodies have shown an unprecedented pace of development and have brought new hope for the treatment of numerous diseases. The bioinformatic tools for modelling antibody structures have become invaluable for antibody engineering and the development of therapeutic antibodies. The antigen-binding site consists of six hypervariable loops, also known as the Complementary Determining Regions (CDR), all of which can be modelled with adequate accuracy, except for one. It remains markedly difficult to model the third CDR loop of the antibody heavy chain. The CDRH3 differs in length, has far greater sequence variability and has such a great structural diversity that modelling it is considerably harder. }}\label{therapeutic-antibodies-have-shown-an-unprecedented-pace-of-development-and-have-brought-new-hope-for-the-treatment-of-numerous-diseases.-the-bioinformatic-tools-for-modelling-antibody-structures-have-become-invaluable-for-antibody-engineering-and-the-development-of-therapeutic-antibodies.-the-antigen-binding-site-consists-of-six-hypervariable-loops-also-known-as-the-complementary-determining-regions-cdr-all-of-which-can-be-modelled-with-adequate-accuracy-except-for-one.-it-remains-markedly-difficult-to-model-the-third-cdr-loop-of-the-antibody-heavy-chain.-the-cdrh3-differs-in-length-has-far-greater-sequence-variability-and-has-such-a-great-structural-diversity-that-modelling-it-is-considerably-harder.}}

\hypertarget{many-sophisticated-approaches-for-antibody-modelling-such-as-the-abymod-software-have-been-developed.-although-such-efforts-have-improved-prediction-accuracy-the-results-for-the-cdrh3-loop-are-still-inconsistent-and-require-further-improvement.-providing-a-confidence-score-for-the-structure-predictions-would-aid-in-differentiating-well-modelled-structures-from-incorrectly-modelled-structures-giving-the-abymod-user-a-clearer-understanding-of-the-generated-model-reliability.}{%
\section{\texorpdfstring{Many sophisticated approaches for antibody
modelling, such as the abYmod software, have been developed. Although
such efforts have improved prediction accuracy the results for the CDRH3
loop are still inconsistent and require further improvement. Providing a
confidence score for the structure predictions would aid in
differentiating well-modelled structures from incorrectly modelled
structures, giving the abYmod user a clearer understanding of the
generated model reliability.
}{Many sophisticated approaches for antibody modelling, such as the abYmod software, have been developed. Although such efforts have improved prediction accuracy the results for the CDRH3 loop are still inconsistent and require further improvement. Providing a confidence score for the structure predictions would aid in differentiating well-modelled structures from incorrectly modelled structures, giving the abYmod user a clearer understanding of the generated model reliability. }}\label{many-sophisticated-approaches-for-antibody-modelling-such-as-the-abymod-software-have-been-developed.-although-such-efforts-have-improved-prediction-accuracy-the-results-for-the-cdrh3-loop-are-still-inconsistent-and-require-further-improvement.-providing-a-confidence-score-for-the-structure-predictions-would-aid-in-differentiating-well-modelled-structures-from-incorrectly-modelled-structures-giving-the-abymod-user-a-clearer-understanding-of-the-generated-model-reliability.}}

\hypertarget{we-present-a-model-quality-predictor-combining-domain-knowledge-with-machine-learning-techniques-to-predict-the-accuracy-of-cdrh3-models-generated-by-abymod.-the-newly-developed-predictor-scored-a-mathews-correlation-coefficient-of-0.63-thus-its-reliability-could-be-demonstrated.}{%
\section{We present a model quality predictor, combining domain
knowledge with machine learning techniques to predict the accuracy of
CDRH3 models generated by abYmod. The newly developed predictor scored a
Mathews Correlation Coefficient of 0.63, thus its reliability could be
demonstrated.}\label{we-present-a-model-quality-predictor-combining-domain-knowledge-with-machine-learning-techniques-to-predict-the-accuracy-of-cdrh3-models-generated-by-abymod.-the-newly-developed-predictor-scored-a-mathews-correlation-coefficient-of-0.63-thus-its-reliability-could-be-demonstrated.}}

\hypertarget{we-conclude-our-predictor-may-prove-useful-for-abymod-users-and-would-be-a-useful-future-addition-to-the-modelling-software-web-interface.}{%
\section{\texorpdfstring{We conclude our predictor may prove useful for
abYmod users and would be a useful future addition to the modelling
software web interface.
}{We conclude our predictor may prove useful for abYmod users and would be a useful future addition to the modelling software web interface. }}\label{we-conclude-our-predictor-may-prove-useful-for-abymod-users-and-would-be-a-useful-future-addition-to-the-modelling-software-web-interface.}}

Antibodies are highly specialized proteins of the immune system that are
produced in response to a foreign substance, called an antigen. A mature
antibody binds a specific antigen with high affinity, while only weakly
interacting with other antigens, or not at all. This high affinity, high
specificity sets it apart from other pharmaceuticals. Furthermore, in
contrast with small drug molecules, antibodies can not only bind
pockets, but also flat, concave or even convex
surfaces\textsuperscript{1}. Their unique characteristics have enabled
researchers to develop efficient antibody drugs for treating cancers,
autoimmune disorders, infectious diseases and many more
\textsuperscript{2}. Their ability to target an immense variety of
antigens allows for endless possibilities in application. The global
market size was valued at USD 130.9 billion in 2020, estimated to grow
223.7 billion by the end of 2025 at a compound annual growth rate of
11.31\% \textsuperscript{3}. Four of the top 10 best-selling drugs in
2020 were monoclonal antibodies \textsuperscript{4}.

In order to rationally design therapeutic antibodies, knowledge of their
structure is essential. The acquired structural information can be used
to increase binding affinity to a target of interest, predicting both
the exact binding site and the antibody stability as well as assessing
immunogenicity \textsuperscript{5}. As experimental structure
determination is very costly and time consuming, computational
predictions of an antibody's structure are used to streamline the
process.

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known as
hypervariable loops. All except one of these loops can be clustered into
a limited number of canonical structures. Therefore, modelling these
loops with adequate accuracy is commonly achievable
\textsuperscript{6,7}. However, the CDR loop 3 of the heavy chain
(CDRH3) has a far greater sequence variability due to the processes of
V(D)J recombination and somatic hyper‐mutation and its structure has
remained unclassifiable \textsuperscript{8}. The variety in structure is
so great, that its structural diversity is remarkable even compared to
other protein loops \textsuperscript{9}. It was found that over 75\% of
CDRH3 loops do not have a sub-Angstrom non-antibody structural
neighbour, as well as that 30\% of CDRH3 loops have a completely unique
structure compared with under 3\% for all other loops on average
\textsuperscript{9}.

Apart from being the most structurally diverse, the H3 loop is also the
most important for antigen binding, being located at the center of the
binding site and forming the most contacts to the antigen
\textsuperscript{10}. In fact, it was demonstrated that differences in
this loop alone were sufficient to enable otherwise identical antibodies
to distinguish between various antigens \textsuperscript{11}.

\includegraphics[width=6.51667in,height=2.55833in]{media/image1.png}
According to the Kabat definition, the CDRH3 loop is made up of the
residues 95-105 (Kabat numbering scheme\textsuperscript{12}) in the
heavy chain, with a potential insertion site at position 100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long (Fig.1).

For shorter loops, a higher prediction accuracy can be achieved than for
longer CDRH3 loops. This was also shown by the Antibody Modelling
Assessments (AMA), two blind contests that required researchers to build
structural models from antibody sequences. The CDR-H3 loop modelling
quality achieved at the contests was on average much lower for loops of
longer lengths \textsuperscript{13,14}.

Several different approaches for generating three-dimensional structure
models from antibody sequences exist such as RosettaAntibody
\textsuperscript{15,16}, ABodyBuilder\textsuperscript{17},
PIGSPro\textsuperscript{18} and abYmod, UCL's in-house software
developed by Prof. Martin. One of the most used methods is
RosettaAntibody, which implements template selection and ab initio
CDR-H3 loop modelling using loop fragments and employing specific angle
restraints which bias the conformational space towards so-called
`kinked' loops\textsuperscript{19,20}. In contrast, ABodyBuilder uses a
database search algorithm (FREAD\textsuperscript{21}) for CDR loop
modelling.

In this project abYmod is used, which is to be found at
\url{http://abymod.abysis.org/}. The program utilizes extensive
canonical class definitions, VH/VL angle prediction and a large database
of loop structures (LoopDB) for CDR-H3 modelling to achieve optimal
results. Upon inputting an antibody sequence, abYmod assigns the
canonical class using a set of key residues \textsuperscript{22}and
where an exact match is not possible, a nearest match is made. Then, the
program identifies the 10 best overall-matching PDB files according to
sequence identity for the light and heavy chain. Of these, the best
template is then identified for each CDR using sequence similarity and
identity. In general, modelling the antibody using the single best
overall-matching template works best and the CDR-specific templates are
used if there is no canonical match. The VH/VL packing angle is then
determined either using machine learning or the chosen template
structure for one of the chains. CDR specific templates, if selected,
are grafted onto the framework. If there is no template of the correct
length for CDR-H3, the loop is built using the LoopDB database,
containing CDRH3-like loops from all proteins. Finally, Gromacs energy
minimization software is used to optimize the model. This method has
proven very effective and preliminary analysis suggests the method
achieves comparable results or outperforms other modelling software (see
results section).

\includegraphics[width=5.78542in,height=3.34583in]{media/image7.png}\includegraphics[width=2.61914in,height=2.05972in]{media/image9.png}
Using these mentioned modelling methods, framework regions can generally
be predicted with great accuracy (with better than 1Å
RMSD\textsuperscript{13}), as one can often find a very similar
structure for the homology modelling process. However, the CDR loops are
not as easily predicted due to their great diversity. If the canonical
conformation of CDR loops 1-5 can be identified, they too can be
modelled rather well, often within 1Å RMSD, for CDR-H3 loops the average
is usually above 3Å \textsuperscript{13}.

\includegraphics[width=3.32639in,height=3.65972in]{media/image10.png}To
our knowledge, ABodyBuilder is the only modelling server that provides
the user with a confidence score for each region (e.g.CDR-H2) of the
antibody model. The given score is the probability that a specific
region (e.g. CDR-H2) will be Thus, it can be used to obtain an expected
RMSD value for a given probability (default 75\%). For the CDR-H3 this
score is calculated as a function of the loop length. The confidence
scorer is described as robust, but less accurate in the case of CDR
loops due to the lack of data.\textsuperscript{17}

Modelling the H3 loop is a hurdle for in silico development of
therapeutic antibodies. Currently, there is no reliable way to determine
how accurate a generated structural model is within the H3 region.
Therefore, the aim of this project will be to produce a user-friendly
predictor of H3 model quality. The predictor will give the user an
RMSD-range in Ångstöms, in which the generated model lies with a high
probability. Making such a confidence score available via the web
interface of the in-house modelling software abYmod is a future goal.
Such a score is not provided by most modelling programs and would thus
be a novel addition. This information can guide the user in the antibody
engineering process. The user has the choice to determine whether the
model is to be used in the intended way, or whether the model should be
re-worked.

\textbf{RESULTS}

predictive power regarding the CDR-H3 loop, the software was tested on a
test-set of antibody structures used in the 2014 and 2011 Antibody
Modelling Assessments \textsuperscript{13,23}. As the results depicted
in \emph{fig}.2 show, abYmod achieves results similar to, or better than
other modelling programs. However, the outliers with very high RMSD
values increase abYmod's RMSD average. The predictor in this work would
aim to identify such outlier models.

The predictive power of any machine learning model is largely dependent
on the quality and size of the dataset it was trained on. As this is a
non-linear, complex, multi-class classification problem, a substantial
amount of data was required. Thus, an extensive, verified dataset of
antibody structures called abYbank/AbDb\textsuperscript{24}, established
by Prof. Andrew Martin, was utilised (1924 non-redundant structures).
The root-mean-square deviation (RMSD) value, a measure of distance
between backbone C-α atoms of superimposed crystal structures and
modelled structures, is calculated (see methods). This metric for model
quality was used to classify models.

\includegraphics[width=3.60764in,height=3.48125in]{media/image14.png}The
full pipeline for creating the final machine learning model that will
predict model quality by giving its RMSD range is summarized in
\emph{fig.3}. The first step of this process is feature-set calculation
using the antibody sequence. The feature set includes many different
attributes linked to sequence, physical characteristics, interactions,
etc. within as well as outside of the loop. There is a plethora of
information to train our classifier on, including packing quality,
protrusion, hydrophobicity, pseudo-energy (see

protrusion, hydrophobicity, pseudo-energy (see methods for details). The
sequence logo (\emph{fig.4}) visualizes amino acid occurrence within the
loop sequence, elements of which can be extracted as features
\textsuperscript{26,27}.

After creating the feature dataset, it is pre-processed (cleaning,
scaling, encoding, see methods for details). Structures with a
resolution below 4Å were removed given their low quality. Identical
whole antibody structures were removed from the dataset, while instances
of different antibodies that matched in loop sequence were not removed.
Fig.5 illustrates that models of some of these structures with the same
loop sequence differ significantly. The few large RMSD ranges may stem
from low resolution, e.g. the highest datapoint contains a structure
with a resolution of 3.00 Å. Residue differences near the loop may also
explain the conformational difference. Some of these structures are
complexed while others are not, which may also affect the loop
structure.

The dataset was also screened for any models which

abYmod created using a template sequence from LoopDB, a database of
CDR-H3 like loops from all proteins. This was not the case for any of
our structures.

The target data (i.e. RMSD values) are then transformed from numerical
values to nominal values so that they can be used for classification. In
order to define these nominal categories, the total RMSD range must be
divided into categories. This is done either by creating uniform classes
i.e. 1-2Å, 2-3Å (the optimal size of which must be determined), etc. or
by creating balanced classes. When creating balanced classes, the upper
and lower thresholds of a category are chosen in such a way that each
class contains an equal number of instances. This approach is chosen to
counteract the skewness of the RMSD distribution (\emph{fig}.2).
However, this was found to negatively affect the final model's
predictive power. Therefore, uniform classes were used.

They are also transformed into a set of binary values according to a
list of RMSD thresholds. This is done so that binary models can be
trained, which will predict the probability e.g. that the model's RMSD
is above 2Å, 2.2Å, 2.4Å, and so on. The number of binary classifiers
incorporated into the first layer

have a great effect on the final model, the general

trend being that the more binary classifiers are used, the better the
nominal prediction.

\textbf{Feature Encoding and Selection}

As some features are in the form of amino acid codes, these must be
encoded before they can be passed to a machine learning model. The
encoding strategy often determines how efficiently the model learns and
how much information can be extracted. Different strategies were
employed to represent amino acids as a numerical value: one-hot
encoding, BLOSUM62 encoding, NLF encoding and a 4-feature physiochemical
encoding method (see methods section). The effect of encoding strategy
on the ML model's performance was evaluated. It was

\includegraphics[width=7.25in,height=3.81111in]{media/image16.png}found
that one-hot encoding was the least effective,

followed by BLOSUM62 and NLF encoding. The physiochemical encoding
strategy was implemented for all models, being the most effective.
However, PCA-3 BLOSUM62, a dimensionality-reduced BLOSUM62 encoding
method achieved comparable results.

Feature selection was conducted to improve the ML model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating model overfitting and can be
responsible for an overall decrease in model performance and stability.
Each additionally inputted feature forces the model to handle a more
complex task, which consumes excess computational power and time and
leads to overfitting of the model.

Our model is trained on different feature sets selected using manual
selection as well as algorithmic selection strategies (see appendix), in
order to determine the most effective feature selection method. None of
the feature selection methods was a best fit for all models. However,
the

After the data is processed, it can be fed into different machine
learning models. Different model types are investigated, as the most
suited model-type has to be heuristically determined. We decided on the
following list, which includes some of the most commonly used
algorithms: logistic regression, linear discriminant analysis, K-nearest
neighbours classifier, decision tree classifier, Gaussian NB, random
forest classifier, support vector machine, probability-based voting
(also known as soft voting) and extreme gradient boosting
(XGBoost)\textsuperscript{28}.

The best model, and its best hyperparameters, are then determined for
each binary RMSD target. The set of binary models outputs a number of
predictions that give the likelihood of the model having an RMSD above
the threshold value of the respective model. These predictions are then
added to the feature set, which a top-layer classifier is then trained
on. Thus, a quasi-voting-system is incorporated into the final
classifier, in which a set of weaker classifiers vote on the model
quality.

\textbf{Hyperparameter Optimization}

In the process of hyperparameter optimization, the configuration of
model parameters which results in best performance is selected. This is
usually a computationally expensive and manual procedure.

In an effort to automate this process, a population is defined for each
model type, so hyperparameter optimization can be conducted
automatically for each model and seamlessly integrated into the full
model creation process. Two different methods for hyperparameter
optimization were tested. The first is a hybrid approach of randomized
search and grid search, the second uses a genetic algorithm for
optimization. The genetic algorithm was found to achieve slightly better
results and was selected for all models.

\textbf{Model Performance}

The overall best final model is composed of several different binary
classifiers, with an extreme gradient boosting (XGBoost) top-layer
nominal classifier. Features were selected using random forest feature
selection. A final MCC value of 0.64 could be achieved.

The classifier predicts whether a model has an RMSD of 0-2Å (`class 0'),
2-4Å (`class 1') or above 4Å (`class 2'). These cut-off values were
selected based on the observation that abYmod generally produces a model
with RMSD below 4Å. Incorrectly modelled structures (such as the ones
seen in Fig 2.) may be identified by screening for `class 2'. If a very
high-quality model is needed one should find this in `class 0'.

\hypertarget{section}{%
\section{}\label{section}}

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

\hypertarget{section-1}{%
\section{}\label{section-1}}

The results suggest that our presented classifier can differentiate
between well-modelled and less well-modelled CDR-H3 loop structures. An
MCC value of 0.63 was achieved, which underlines this ability for
accurate discrimination. Different methods for data pre-processing,
feature encoding, feature selection and hyperparameter optimization were
tested.

Feature encoding methods that were very high-dimensional
(one-hot-encoding, BLOSUM62, NLF) were found to be unfavorable.
Dimensionality reduction methods (Principal Component Analysis (PCA),
Independent Component Analysis (ICA), projection-based methods e.g.
t-SNE) were used on BLOSUM62 encoded matrices, which lead to significant
improvement. However, a physiochemical encoding strategy was most
effective.

The selection of features incorporated in the training set seemed to be
most important for effective learning. A multitude of methods were
tested. No one-fit-for-all method for the different models could be
found. However, for our top-layer classifier in our final model
recursive feature elimination worked best.

A set of commonly used machine learning algorithms were tested, and the
best ML models were incorporated into the final ensemble model. A
stacked model approach (consisting of 23 binary classifiers and a single
top-layer nominal classifier) was shown to outperform single ML models.

An MCC value of 0.63 was achieved for a classifier predicting whether an
input-model has an RMSD value below 2Å, 2Å-4Å or above 4Å.

Further research may enable additional optimization of the classifier
making it more reliable.

Given that abYbank/AbDb is soon to be expanded by an additional ca. 2000
structures, classifier performance may be improved by a larger dataset.

At a later stage the model could provide an abYmod user with information
on the generated model's quality in the CDR-H3 region via a
web-interface.

It is conceivable that the described predictor may also be incorporated
in the antibody modelling process as a low-quality filter in the future,
flagging certain structures for re-modelling.

In a future research project residue patterns in correlation with RMSD
may be analyzed. Possibly, one might identify certain sequence patterns
that make accurate modelling with abYmod more difficult. Furthermore,
separate classifiers according to loop length can be built. Given that
loop length is the most important determinant of model quality, this
approach may yield some insight into the challenges of modelling shorter
vs longer loops.

One could also conduct an analysis of the predictor's behaviour when
abYmod is forced to use LoopDB-based modelling. This might shed light on
whether the ML model presented in this paper is biased towards abYmod's
used source of template structures. It would also give an indication of
how well the predictor would work in combination with other modelling
software.

\hypertarget{experimental-procedures}{%
\section{EXPERIMENTAL PROCEDURES}\label{experimental-procedures}}

\hypertarget{section-2}{%
\section{}\label{section-2}}

\textbf{Computing}

All machine learning and, feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training models, the
Yellowbrick\textsuperscript{29} library was utilized for visualization.
All code is available at
\url{https://github.com/LilianDenzler/qualiloop} .

The code was run under CentOS 7 on an 8-core virtual machine on an Intel
Xeon 4208 CPU with 16Gig RAM.

\begin{longtable}[]{@{}lll@{}}
\toprule
Feature Name & Description & Method of Calculation\tabularnewline
\midrule
\endhead
Sequence & Amino acid sequence of the CDR-H3 loop. & Sequence is given
in one-letter amino acid codes.\tabularnewline
Length & Number of residues in the CDRH3-loop, which is located at
residues H95-H102. & The number of residues are counted.\tabularnewline
Sequence identity & Sequence identity of selected template (SeqA) with
input loop sequence (SeqB)) is determined after sequence alignment.
Calculated by abYmod during modelling. &
\(\text{Identity}\left( \text{Seq}A,SeqB \right) = 100\%*\frac{\text{identical\ residues}}{length(alignment)}\)\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Sequence similarity\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Sequence similarity of selected template (SeqA) with input loop sequence
(SeqB) is determined after sequence alignment. Calculated by abYmod
during modelling.

Similar residues are residues that have undergone conservative
substitution.\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\[\text{Simialrity}\left( \text{Seq}A,SeqB \right) = 100\%*\frac{identical\ residues + similar\ residues}{length(alignment)}\]\strut
\end{minipage}\tabularnewline
Loop protrusion & Distance of loop residue further away from the loop
base. & Geometrical calculations, see \emph{fig. 8}\tabularnewline
Protruding residue & The amino acid code of the most protruding loop
residue & Using the previously determined point furthest away from the
loop base, the residue at this coordinate is determined and given as a
one-letter amino acid code.\tabularnewline
Charge & Total charge of the loop & Sum of charges of all residues in
loop\tabularnewline
Charge difference & Difference in total charge compared to template
sequence & Difference between the two summed charges\tabularnewline
Hydrophobicity & Mean of hydrophobicity values of loop & Based Eisenberg
consensus values\textsuperscript{30}\tabularnewline
Hydrophobicity difference & Sum of absolute differences between loop
sequence and template loop & Based Eisenberg consensus
values\textsuperscript{30}\tabularnewline
Accessibility & Total and average accessibility for the loop. &
Lee-Richards method \textsuperscript{31} implemented using the pdbsolv
method from the BiopTools library \textsuperscript{32}\tabularnewline
Side-chain Accessibility & Total and average side-chain accessibility
for the loop. & Lee-Richards method \textsuperscript{31} implemented
using the pdbsolv method from the BiopTools library
\textsuperscript{32}\tabularnewline
Relative Accessibility & Total and average relative accessibility for
the loop. & Lee-Richards method \textsuperscript{31} implemented using
the pdbsolv method from the BiopTools library
\textsuperscript{32}\tabularnewline
Relative side-chain accessibility & Total and average relative
side-chain accessibility for the loop. & Lee-Richards method
\textsuperscript{31} implemented using the pdbsolv method from the
BiopTools library \textsuperscript{32}\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Happiness\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Happiness score, taking accessibility and hydrophobicity into account.
If a residue is `happy' it will not be a buried hydrophilic or a surface
hydrophobic residue.\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Hydrophobicity values (see above) are normalized to a range of -1 to +1.
Mean accessibility values are calculated as above.

If Hydrophobicity of loop is \textless0:

\[Happiness = 1 + (Hydrophobicity*(1 - Accessibility)\]

Otherwise:

\[Happiness = 1 - (Hydrophobicity*Accessibility)\]\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\textbf{Feature Calculations}

\begin{longtable}[]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.30\columnwidth}\raggedright
Nr. Of contacts\strut
\end{minipage} & \begin{minipage}[b]{0.30\columnwidth}\raggedright
Nr of contacts made by the residue of the loop within a range of 3.5Å.
Includes mainchain as well as sidechain atoms. Contacts made with
residue within and outside of the loop are counted separately and as
total.

The ratio of inside vs outside is also calculated.\strut
\end{minipage} & \begin{minipage}[b]{0.30\columnwidth}\raggedright
Modified version of the rangecontacts method in the BiopTools library
\textsuperscript{32}.\strut
\end{minipage}\tabularnewline
\midrule
\endhead
Energy & Potential energy of the model. & Calculated by
Gromacs\textsuperscript{33} during energy minimization step in abYmod
modelling.\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Lowest BLOSUM 62 Scoring Residue Pair\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Each possible residue pair in the CDR-H3 loop is scored by their BLOSUM
62 score. The lowest scoring pair's BLOSUM62 value will be combined with
their residue separation to form the metric.\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
With separation being the number of residues between the worst residue
pair, and the worst score being the lowest BLOSUM62 score achieved by a
residue pair, the metric is calculated as follows:

\[WorstBLOSUM = - \log_{2}\left( \text{separation} \right)*worst\ score\ \]\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\includegraphics[width=6.47569in,height=2.2744in]{media/image18.png}

\textbf{Data Pre-Processing and Preparation}

{\emph{Handling Null Values and Duplicates}\textbf{: }}The dataset
containing target RMSD values, and the calculated features was screened
for null values. If a feature column contained more than 5\% null
values, it was dropped (none removed). Rows that contained any null
values were removed from the dataset (11 rows in total).

{Duplicate Screening}: Using AbDb's redundancy information it was
ensured that no antibodies were present in the dataset more than once.
The dataset is additionally screened for duplicate instances.

{Scaling:} Normalization and Standardization are tested as scaling
methods. Both approaches are greatly influenced by outliers, and such
datapoints are ideally removed for optimal scaling. Here we define
outliers as datapoints that lie over 1.5 times the interquartile range
(IQR) below the first quartile or above the third

quartile. The IQR is defined as the range between quartile 1, i.e. the
median of the lower half of the data, and quartile 3, i.e. the median of
the upper half of the

data. However, across all features there are a total of 632 outlier
values and removing such a large number of datapoints is not a viable
option. A robust scaler was also used, which uses statistics that are
robust to outliers. The median is set to zero and numerical features are
scaled to the interquartile range.

{BLOSUM 62 encoding:} The BLOSUM62 matrix reflects the frequencies of
amino acid substitutions within a locally aligned, conserved regions of
proteins with at least 62\% similarity. Each amino acid is represented
by a row (or column) of the BLOSUM62 matrix. Dimensionality reduction
techniques are employed: Principal Component Analysis (PCA), Independent
Component Analysis (ICA), projection-based methods (t-SNE, Isomap).
Three components were used as features.

{Physiochemical Feature Encoding\emph{:}} In a paper by L. Nanni and
A.Lumini a new encoding technique is

presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction. A vector of
length 19 is produced for each amino acid\textsuperscript{34}.

In a paper on designing a neural network for predicting the packing
angle of the light and heavy variable chain of an antibodies, A.C.R.
Martin and K.R. Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector\textsuperscript{25}. The
amino acid properties used are 1) the total number of side-chain atoms,
2) the number of side-chain atoms in the shortest path from Cα to the
most distal atom, 3) the Eisenberg consensus
hydrophobicity\textsuperscript{30}, 4) the charge.

\textbf{Dataset-splitting}: The final model was evaluated using a test
set, separated from the training set at the start in a 30/70 split
(lock-box principle) \textsuperscript{35}. The performance of all
individual sub-models of the first layer are determined using stratified
K-folds cross-validation (k=10) as the dataset is imbalanced, being
skewed towards lower RMSD values. The method is differentiable from
K-folds cross validation as it uses stratified sampling instead of
random sampling. This ensures each class is represented, as the
percentage of samples for each class are preserved.

A validation set, usually used for testing during the optimization stage
will be omitted in favour of stratified K-folds cross-validation (k=10)
\textsuperscript{36,37}.

\textbf{Model Assessment:} Model assessment must be considered at two
levels as performance metrics of binary and multi-class classifiers are
calculated differently and must thus be considered separately. The
Mathews Correlation Coefficient (MCC) \textsuperscript{38} is deemed the
most informative, taking the ratios of the four confusion matrix
categories into account \textsuperscript{39}and is thus more reliable
than the F1 score and accuracy . It is also consistent for both binary
and multi-class problems and therefore well suited for our purpose.

\hypertarget{section-3}{%
\section{}\label{section-3}}

\hypertarget{section-4}{%
\section{\texorpdfstring{\protect\includegraphics[width=3.66667in,height=3.35694in]{media/image19.png}}{}}\label{section-4}}

\hypertarget{section-5}{%
\section{}\label{section-5}}

\hypertarget{section-6}{%
\section{}\label{section-6}}

\hypertarget{section-7}{%
\section{}\label{section-7}}

\hypertarget{section-8}{%
\section{}\label{section-8}}

\hypertarget{section-9}{%
\section{}\label{section-9}}

\hypertarget{section-10}{%
\section{}\label{section-10}}

\hypertarget{section-11}{%
\section{}\label{section-11}}

\hypertarget{section-12}{%
\section{}\label{section-12}}

\hypertarget{section-13}{%
\section{}\label{section-13}}

\hypertarget{section-14}{%
\section{}\label{section-14}}

\hypertarget{section-15}{%
\section{}\label{section-15}}

\hypertarget{section-16}{%
\section{}\label{section-16}}

\hypertarget{section-17}{%
\section{}\label{section-17}}

\hypertarget{section-18}{%
\section{}\label{section-18}}

\hypertarget{section-19}{%
\section{}\label{section-19}}

\hypertarget{section-20}{%
\section{}\label{section-20}}

\hypertarget{section-21}{%
\section{}\label{section-21}}

\hypertarget{section-22}{%
\section{}\label{section-22}}

\hypertarget{section-23}{%
\section{}\label{section-23}}

\hypertarget{section-24}{%
\section{}\label{section-24}}

\hypertarget{section-25}{%
\section{}\label{section-25}}

\hypertarget{references}{%
\section{REFERENCES}\label{references}}

\hypertarget{section-26}{%
\section{}\label{section-26}}

\hypertarget{charles-a-janeway-j.-travers-p.-walport-m.-shlomchik-m.-j.-the-interaction-of-the-antibody-molecule-with-specific-antigen.-immunobiology-the-immune-system-in-health-and-disease.-garland-science-2001.2.-lu-r.-m.-et-al.-development-of-therapeutic-antibodies-for-the-treatment-of-diseases.-journal-of-biomedical-science-vol.-27-130-2020.3.-antibodies-market-size-share-trends-growth-forecast-2020-to-2025.-httpswww.marketdataforecast.commarket-reportsantibodies-market.4.-urquhart-l.-top-companies-and-drugs-by-sales-in-2020.-nat.-rev.-drug-discov.-2021-doi10.1038d41573-021-00050-6.5.-abhinandan-k.-r.-martin-a.-c.-r.-analyzing-the-degree-of-humanness-of-antibody-sequences.-j.-mol.-biol.-369-852862-2007.6.-north-b.-lehmann-a.-dunbrack-r.-l.-a-new-clustering-of-antibody-cdr-loop-conformations.-j.-mol.-biol.-406-228256-2011.7.-weitzner-b.-d.-dunbrack-r.-l.-gray-j.-j.-the-origin-of-cdr-h3-structural-diversity.-structure-23-302311-2015.8.-finn-j.-a.-et-al.-improving-loop-modeling-of-the-antibody-complementarity-determining-region-3-using-knowledge-based-restraints.-plos-one-11-e0154811-2016.9.-regep-c.-georges-g.-shi-j.-popovic-b.-deane-c.-m.-the-h3-loop-of-antibodies-shows-unique-structural-characteristics.-proteins-struct.-funct.-bioinforma.-85-13111318-2017.10.-maccallum-r.-m.-martin-a.-c.-r.-thornton-j.-m.-antibody-antigen-interactions-contact-analysis-and-binding-site-topography.-j.-mol.-biol.-262-732745-1996.11.-xu-j.-l.-davis-m.-m.-diversity-in-the-cdr3-region-of-vh-is-sufficient-for-most-antibody-specificities.-immunity-13-3745-2000.12.-kabat-e.-wu-t.-te-perry-h.-foeller-c.-gottesman-k.-sequences-of-proteins-of-immunological-interest.-1992.13.-almagro-j.-c.-et-al.-second-antibody-modeling-assessment-ama-ii.-proteins-structure-function-and-bioinformatics-vol.-82-15531562-2014.14.-almagro-j.-c.-et-al.-antibody-modeling-assessment.-proteins-struct.-funct.-bioinforma.-79-30503066-2011.15.-sircar-a.-kim-e.-t.-gray-j.-j.-rosettaantibody-antibody-variable-region-homology-modeling-server.-nucleic-acids-res.-37-w474w479-2009.16.-sivasubramanian-a.-sircar-a.-chaudhury-s.-gray-j.-j.-toward-high-resolution-homology-modeling-of-antibody-f-v-regions-and-application-to-antibody-antigen-docking.-proteins-struct.-funct.-bioinforma.-74-497514-2009.17.-leem-j.-dunbar-j.-georges-g.-shi-j.-deane-c.-m.-abodybuilder-automated-antibody-structure-prediction-with-datadriven-accuracy-estimation.-mabs-8-12591268-2016.18.-lepore-r.-olimpieri-p.-p.-messih-m.-a.-tramontano-a.-pigspro-prediction-of-immunoglobulin-structures-v2.-nucleic-acids-res.-45-w17w23-2017.19.-schoeder-c.-t.-et-al.-modeling-immunity-with-rosetta-methods-for-antibody-and-antigen-design.-60.20.-weitzner-b.-d.-gray-j.-j.-accurate-structure-prediction-of-cdr-h3-loops-enabled-by-a-novel-structure-based-c-terminal-constraint.-j.-immunol.-198-505515-2017.21.-choi-y.-deane-c.-m.-fread-revisited-accurate-loop-structure-prediction-using-a-database-search-algorithm.-proteins-struct.-funct.-bioinforma.-78-14311440-2010.22.-martin-a.-c.-r.-thornton-j.-m.-structural-families-in-loops-of-homologous-proteins-automatic-classification-modelling-and-application-to-antibodies.-j.-mol.-biol.-263-800815-1996.23.-almagro-j.-c.-et-al.-antibody-modeling-assessment.-proteins-struct.-funct.-bioinforma.-79-30503066-2011.24.-ferdous-s.-martin-a.-c.-r.-abdb-antibody-structure-databasea-database-of-pdb-derived-antibody-structures.-database-2018-2018.25.-abhinandan-k.-r.-martin-a.-c.-r.-analysis-and-prediction-of-vhvl-packing-in-antibodies.-protein-eng.-des.-sel.-23-689697-2010.26.-thomsen-m.-c.-f.-nielsen-m.-seq2logo-a-method-for-construction-and-visualization-of-amino-acid-binding-motifs-and-sequence-profiles-including-sequence-weighting-pseudo-counts-and-two-sided-representation-of-amino-acid-enrichment-and-depletion.-nucleic-acids-res.-40-2012.27.-shaner-m.-c.-blairs-i.-m.-schneider-t.-d.-sequence-logos-a-powerful-yet-simple-tool.28.-chen-t.-guestrin-c.-xgboost-a-scalable-tree-boosting-system.-in-proceedings-of-the-acm-sigkdd-international-conference-on-knowledge-discovery-and-data-mining-vols-13-17-august-2016-785794-association-for-computing-machinery-2016.29.-bengfort-b.-et-al.-yellowbrick-v1.3.-2021-doi10.5281zenodo.4525724.30.-eisenberg-d.-weiss-r.-m.-terwilliger-t.-c.-wilcox-w.-hydrophobic-moments-and-protein-structure.-in-faraday-symposia-of-the-chemical-society-vol.-17-109120-the-royal-society-of-chemistry-1982.31.-lee-b.-richards-f.-m.-the-interpretation-of-protein-structures-estimation-of-static-accessibility.-j.-mol.-biol.-55-379-in4-1971.32.-porter-c.-t.-martin-a.-c.-r.-bioplib-and-bioptools---a-c-programming-library-and-toolset-for-manipulating-protein-structure.-bioinformatics-31-40174019-2015.33.-van-der-spoel-d.-et-al.-gromacs-fast-flexible-and-free.-j.-comput.-chem.-26-17011718-2005.34.-nanni-l.-lumini-a.-a-new-encoding-technique-for-peptide-classification.-expert-syst.-appl.-38-31853191-2011.35.-wong-t.-t.-performance-evaluation-of-classification-algorithms-by-k-fold-and-leave-one-out-cross-validation.-pattern-recognit.-48-28392846-2015.36.-krstajic-d.-buturovic-l.-j.-leahy-d.-e.-thomas-s.-cross-validation-pitfalls-when-selecting-and-assessing-regression-and-classification-models.-j.-cheminform.-6-10-2014.37.-kohavi-r.-a-study-of-cross-validation-and-bootstrap-for-accuracy-estimation-and-model-selection.-httprobotics.stanford.eduronnyk-1995.38.-structure-b.-m.-b.-et-b.-a.-bba-p.-1975-undefined.-comparison-of-the-predicted-and-observed-secondary-structure-of-t4-phage-lysozyme.-elsevier.39.-chicco-d.-ten-quick-tips-for-machine-learning-in-computational-biology.-biodata-mining-vol.-10-35-2017.40.-morris-j.-the-precise-effect-of-multicollinearity-on-classification-prediction-ridge-regression-view-project-curriculum-based-vam-view-project.-httpswww.researchgate.netpublication336561703-2014.41.-handling-multi-collinearity-in-ml-models-by-vishwa-pardeshi-towards-data-science.-httpstowardsdatascience.comhandling-multi-collinearity-6579eb99fd81.42.-jin-lee-c.-park-c.-s.-seok-kim-j.-baek-j.-g.-a-study-on-improving-classification-performance-for-manufacturing-process-data-with-multicollinearity-and-imbalanced-distribution.-41-2533-2015.43.-box-p.-o.-van-der-maaten-l.-postma-e.-van-den-herik-j.-tilburg-centre-for-creative-computing-dimensionality-reduction-a-comparative-review-dimensionality-reduction-a-comparative-review.-httpwww.uvt.nlticc-2009.44.-vlachos-m.-et-al.-non-linear-dimensionality-reduction-techniques-for-classification-and-visualization.-2002.45.-sorzano-c.-o.-s.-vargas-j.-pascual-montano-a.-a-survey-of-dimensionality-reduction-techniques.46.-rifai-s.-vincent-p.-muller-x.-glorot-x.-bengio-y.-contractive-auto-encoders-explicit-invariance-during-feature-extraction.-2011.47.-chicco-d.-jurman-g.-the-advantages-of-the-matthews-correlation-coefficient-mcc-over-f1-score-and-accuracy-in-binary-classification-evaluation.-bmc-genomics-21-6-2020.48.-brown-j.-b.-classifiers-and-their-metrics-quantified.-mol.-inform.-37-1700127-2018.49.-delgado-r.-tibau-x.-a.-why-cohens-kappa-should-be-avoided-as-performance-measure-in-classification.-plos-one-14-e0222916-2019.50.-gorodkin-j.-comparing-two-k-category-assignments-by-a-k-category-correlation-coefficient.-comput.-biol.-chem.-28-367374-2004.51.-wong-w.-k.-leem-j.-deane-c.-m.-comparative-analysis-of-the-cdr-loops-of-antigen-receptors-protein-structure-prediction-view-project-comparative-analysis-of-the-cdr-loops-of-antigen.-doi10.1101709840.}{%
\section{\texorpdfstring{1. Charles A Janeway, J., Travers, P., Walport,
M. \& Shlomchik, M. J. \emph{The interaction of the antibody molecule
with specific antigen}. \emph{Immunobiology: The Immune System in Health
and Disease.} (Garland Science, 2001).2. Lu, R. M. \emph{et al.}
Development of therapeutic antibodies for the treatment of diseases.
\emph{Journal of Biomedical Science} vol. 27 1--30 (2020).3. Antibodies
Market Size, Share Trends, Growth, Forecast \textbar{} 2020 to 2025.
https://www.marketdataforecast.com/market-reports/antibodies-market.4.
Urquhart, L. Top companies and drugs by sales in 2020. \emph{Nat. Rev.
Drug Discov.} (2021) doi:10.1038/d41573-021-00050-6.5. Abhinandan, K. R.
\& Martin, A. C. R. Analyzing the `Degree of Humanness' of Antibody
Sequences. \emph{J. Mol. Biol.} \textbf{369}, 852--862 (2007).6. North,
B., Lehmann, A. \& Dunbrack, R. L. A new clustering of antibody CDR loop
conformations. \emph{J. Mol. Biol.} \textbf{406}, 228--256 (2011).7.
Weitzner, B. D., Dunbrack, R. L. \& Gray, J. J. The origin of CDR H3
structural diversity. \emph{Structure} \textbf{23}, 302--311 (2015).8.
Finn, J. A. \emph{et al.} Improving Loop Modeling of the Antibody
Complementarity-Determining Region 3 Using Knowledge-Based Restraints.
\emph{PLoS One} \textbf{11}, e0154811 (2016).9. Regep, C., Georges, G.,
Shi, J., Popovic, B. \& Deane, C. M. The H3 loop of antibodies shows
unique structural characteristics. \emph{Proteins Struct. Funct.
Bioinforma.} \textbf{85}, 1311--1318 (2017).10. MacCallum, R. M.,
Martin, A. C. R. \& Thornton, J. M. Antibody-antigen interactions:
Contact analysis and binding site topography. \emph{J. Mol. Biol.}
\textbf{262}, 732--745 (1996).11. Xu, J. L. \& Davis, M. M. Diversity in
the CDR3 region of V(H) is sufficient for most antibody specificities.
\emph{Immunity} \textbf{13}, 37--45 (2000).12. Kabat, E., Wu, T. Te,
Perry, H., Foeller, C. \& Gottesman, K. Sequences of proteins of
immunological interest. (1992).13. Almagro, J. C. \emph{et al.} Second
Antibody Modeling Assessment (AMA-II). \emph{Proteins: Structure,
Function and Bioinformatics} vol. 82 1553--1562 (2014).14. Almagro, J.
C. \emph{et al.} Antibody modeling assessment. \emph{Proteins Struct.
Funct. Bioinforma.} \textbf{79}, 3050--3066 (2011).15. Sircar, A., Kim,
E. T. \& Gray, J. J. RosettaAntibody: Antibody variable region homology
modeling server. \emph{Nucleic Acids Res.} \textbf{37}, W474--W479
(2009).16. Sivasubramanian, A., Sircar, A., Chaudhury, S. \& Gray, J. J.
Toward high-resolution homology modeling of antibody F \textsubscript{v}
regions and application to antibody-antigen docking. \emph{Proteins
Struct. Funct. Bioinforma.} \textbf{74}, 497--514 (2009).17. Leem, J.,
Dunbar, J., Georges, G., Shi, J. \& Deane, C. M. ABodyBuilder: Automated
antibody structure prediction with data--driven accuracy estimation.
\emph{MAbs} \textbf{8}, 1259--1268 (2016).18. Lepore, R., Olimpieri, P.
P., Messih, M. A. \& Tramontano, A. PIGSPro: Prediction of
immunoGlobulin structures v2. \emph{Nucleic Acids Res.} \textbf{45},
W17--W23 (2017).19. Schoeder, C. T. \emph{et al.} Modeling Immunity with
Rosetta: Methods for Antibody and Antigen Design. \textbf{60},.20.
Weitzner, B. D. \& Gray, J. J. Accurate Structure Prediction of CDR H3
Loops Enabled by a Novel Structure-Based C-Terminal Constraint. \emph{J.
Immunol.} \textbf{198}, 505--515 (2017).21. Choi, Y. \& Deane, C. M.
FREAD revisited: Accurate loop structure prediction using a database
search algorithm. \emph{Proteins Struct. Funct. Bioinforma.}
\textbf{78}, 1431--1440 (2010).22. Martin, A. C. R. \& Thornton, J. M.
Structural families in loops of homologous proteins: Automatic
classification, modelling and application to antibodies. \emph{J. Mol.
Biol.} \textbf{263}, 800--815 (1996).23. Almagro, J. C. \emph{et al.}
Antibody modeling assessment. \emph{Proteins Struct. Funct. Bioinforma.}
\textbf{79}, 3050--3066 (2011).24. Ferdous, S. \& Martin, A. C. R. AbDb:
antibody structure database---a database of PDB-derived antibody
structures. \emph{Database} \textbf{2018}, (2018).25. Abhinandan, K. R.
\& Martin, A. C. R. Analysis and prediction of VH/VL packing in
antibodies. \emph{Protein Eng. Des. Sel.} \textbf{23}, 689--697
(2010).26. Thomsen, M. C. F. \& Nielsen, M. Seq2Logo: A method for
construction and visualization of amino acid binding motifs and sequence
profiles including sequence weighting, pseudo counts and two-sided
representation of amino acid enrichment and depletion. \emph{Nucleic
Acids Res.} \textbf{40}, (2012).27. Shaner, M. C., Blair's, I. M. \&
Schneider, T. D. \emph{Sequence Logos: A Powerful, Yet Simple, Tool}.28.
Chen, T. \& Guestrin, C. XGBoost: A scalable tree boosting system. in
\emph{Proceedings of the ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining} vols 13-17-August-2016 785--794
(Association for Computing Machinery, 2016).29. Bengfort, B. \emph{et
al.} Yellowbrick v1.3. (2021) doi:10.5281/ZENODO.4525724.30. Eisenberg,
D., Weiss, R. M., Terwilliger, T. C. \& Wilcox, W. Hydrophobic moments
and protein structure. in \emph{Faraday Symposia of the Chemical
Society} vol. 17 109--120 (The Royal Society of Chemistry, 1982).31.
Lee, B. \& Richards, F. M. The interpretation of protein structures:
Estimation of static accessibility. \emph{J. Mol. Biol.} \textbf{55},
379-IN4 (1971).32. Porter, C. T. \& Martin, A. C. R. BiopLib and
BiopTools - A C programming library and toolset for manipulating protein
structure. \emph{Bioinformatics} \textbf{31}, 4017--4019 (2015).33. Van
Der Spoel, D. \emph{et al.} GROMACS: Fast, flexible, and free. \emph{J.
Comput. Chem.} \textbf{26}, 1701--1718 (2005).34. Nanni, L. \& Lumini,
A. A new encoding technique for peptide classification. \emph{Expert
Syst. Appl.} \textbf{38}, 3185--3191 (2011).35. Wong, T. T. Performance
evaluation of classification algorithms by k-fold and leave-one-out
cross validation. \emph{Pattern Recognit.} \textbf{48}, 2839--2846
(2015).36. Krstajic, D., Buturovic, L. J., Leahy, D. E. \& Thomas, S.
Cross-validation pitfalls when selecting and assessing regression and
classification models. \emph{J. Cheminform.} \textbf{6}, 10 (2014).37.
Kohavi, R. \emph{A Study of Cross-Validation and Bootstrap for Accuracy
Estimation and Model Selection}.
http://robotics.stanford.edu/\textasciitilde ronnyk (1995).38.
Structure, B. M.-B. et B. A. (BBA)-P. \& 1975, undefined. Comparison of
the predicted and observed secondary structure of T4 phage lysozyme.
\emph{Elsevier}.39. Chicco, D. Ten quick tips for machine learning in
computational biology. \emph{BioData Mining} vol. 10 35 (2017).40.
Morris, J. \emph{The Precise Effect of Multicollinearity on
Classification Prediction Ridge Regression View project Curriculum-Based
VAM View project}. https://www.researchgate.net/publication/336561703
(2014).41. Handling Multi-Collinearity in ML Models \textbar{} by Vishwa
Pardeshi \textbar{} Towards Data Science.
https://towardsdatascience.com/handling-multi-collinearity-6579eb99fd81.42.
Jin Lee, C., Park, C.-S., Seok Kim, J. \& Baek, J.-G. A Study on
Improving Classification Performance for Manufacturing Process Data with
Multicollinearity and Imbalanced Distribution. \textbf{41}, 25--33
(2015).43. Box, P. O., Van Der Maaten, L., Postma, E. \& Van Den Herik,
J. \emph{Tilburg centre for Creative Computing Dimensionality Reduction:
A Comparative Review Dimensionality Reduction: A Comparative Review}.
http://www.uvt.nl/ticc (2009).44. Vlachos, M. \emph{et al.}
\emph{Non-Linear Dimensionality Reduction Techniques for Classification
and Visualization}. (2002).45. Sorzano, C. O. S., Vargas, J. \&
Pascual-Montano, A. \emph{A survey of dimensionality reduction
techniques}.46. Rifai, S., Vincent, P., Muller, X., Glorot, X. \&
Bengio, Y. \emph{Contractive Auto-Encoders: Explicit Invariance During
Feature Extraction}. (2011).47. Chicco, D. \& Jurman, G. The advantages
of the Matthews correlation coefficient (MCC) over F1 score and accuracy
in binary classification evaluation. \emph{BMC Genomics} \textbf{21}, 6
(2020).48. Brown, J. B. Classifiers and their Metrics Quantified.
\emph{Mol. Inform.} \textbf{37}, 1700127 (2018).49. Delgado, R. \&
Tibau, X. A. Why Cohen's Kappa should be avoided as performance measure
in classification. \emph{PLoS One} \textbf{14}, e0222916 (2019).50.
Gorodkin, J. Comparing two K-category assignments by a K-category
correlation coefficient. \emph{Comput. Biol. Chem.} \textbf{28},
367--374 (2004).51. Wong, W. K., Leem, J. \& Deane, C. M. Comparative
analysis of the CDR loops of antigen receptors Protein structure
prediction View project Comparative analysis of the CDR loops of
antigen.
doi:10.1101/709840.}{1. Charles A Janeway, J., Travers, P., Walport, M. \& Shlomchik, M. J. The interaction of the antibody molecule with specific antigen. Immunobiology: The Immune System in Health and Disease. (Garland Science, 2001).2. Lu, R. M. et al. Development of therapeutic antibodies for the treatment of diseases. Journal of Biomedical Science vol. 27 1--30 (2020).3. Antibodies Market Size, Share Trends, Growth, Forecast \textbar{} 2020 to 2025. https://www.marketdataforecast.com/market-reports/antibodies-market.4. Urquhart, L. Top companies and drugs by sales in 2020. Nat. Rev. Drug Discov. (2021) doi:10.1038/d41573-021-00050-6.5. Abhinandan, K. R. \& Martin, A. C. R. Analyzing the `Degree of Humanness' of Antibody Sequences. J. Mol. Biol. 369, 852--862 (2007).6. North, B., Lehmann, A. \& Dunbrack, R. L. A new clustering of antibody CDR loop conformations. J. Mol. Biol. 406, 228--256 (2011).7. Weitzner, B. D., Dunbrack, R. L. \& Gray, J. J. The origin of CDR H3 structural diversity. Structure 23, 302--311 (2015).8. Finn, J. A. et al. Improving Loop Modeling of the Antibody Complementarity-Determining Region 3 Using Knowledge-Based Restraints. PLoS One 11, e0154811 (2016).9. Regep, C., Georges, G., Shi, J., Popovic, B. \& Deane, C. M. The H3 loop of antibodies shows unique structural characteristics. Proteins Struct. Funct. Bioinforma. 85, 1311--1318 (2017).10. MacCallum, R. M., Martin, A. C. R. \& Thornton, J. M. Antibody-antigen interactions: Contact analysis and binding site topography. J. Mol. Biol. 262, 732--745 (1996).11. Xu, J. L. \& Davis, M. M. Diversity in the CDR3 region of V(H) is sufficient for most antibody specificities. Immunity 13, 37--45 (2000).12. Kabat, E., Wu, T. Te, Perry, H., Foeller, C. \& Gottesman, K. Sequences of proteins of immunological interest. (1992).13. Almagro, J. C. et al. Second Antibody Modeling Assessment (AMA-II). Proteins: Structure, Function and Bioinformatics vol. 82 1553--1562 (2014).14. Almagro, J. C. et al. Antibody modeling assessment. Proteins Struct. Funct. Bioinforma. 79, 3050--3066 (2011).15. Sircar, A., Kim, E. T. \& Gray, J. J. RosettaAntibody: Antibody variable region homology modeling server. Nucleic Acids Res. 37, W474--W479 (2009).16. Sivasubramanian, A., Sircar, A., Chaudhury, S. \& Gray, J. J. Toward high-resolution homology modeling of antibody F v regions and application to antibody-antigen docking. Proteins Struct. Funct. Bioinforma. 74, 497--514 (2009).17. Leem, J., Dunbar, J., Georges, G., Shi, J. \& Deane, C. M. ABodyBuilder: Automated antibody structure prediction with data--driven accuracy estimation. MAbs 8, 1259--1268 (2016).18. Lepore, R., Olimpieri, P. P., Messih, M. A. \& Tramontano, A. PIGSPro: Prediction of immunoGlobulin structures v2. Nucleic Acids Res. 45, W17--W23 (2017).19. Schoeder, C. T. et al. Modeling Immunity with Rosetta: Methods for Antibody and Antigen Design. 60,.20. Weitzner, B. D. \& Gray, J. J. Accurate Structure Prediction of CDR H3 Loops Enabled by a Novel Structure-Based C-Terminal Constraint. J. Immunol. 198, 505--515 (2017).21. Choi, Y. \& Deane, C. M. FREAD revisited: Accurate loop structure prediction using a database search algorithm. Proteins Struct. Funct. Bioinforma. 78, 1431--1440 (2010).22. Martin, A. C. R. \& Thornton, J. M. Structural families in loops of homologous proteins: Automatic classification, modelling and application to antibodies. J. Mol. Biol. 263, 800--815 (1996).23. Almagro, J. C. et al. Antibody modeling assessment. Proteins Struct. Funct. Bioinforma. 79, 3050--3066 (2011).24. Ferdous, S. \& Martin, A. C. R. AbDb: antibody structure database---a database of PDB-derived antibody structures. Database 2018, (2018).25. Abhinandan, K. R. \& Martin, A. C. R. Analysis and prediction of VH/VL packing in antibodies. Protein Eng. Des. Sel. 23, 689--697 (2010).26. Thomsen, M. C. F. \& Nielsen, M. Seq2Logo: A method for construction and visualization of amino acid binding motifs and sequence profiles including sequence weighting, pseudo counts and two-sided representation of amino acid enrichment and depletion. Nucleic Acids Res. 40, (2012).27. Shaner, M. C., Blair's, I. M. \& Schneider, T. D. Sequence Logos: A Powerful, Yet Simple, Tool.28. Chen, T. \& Guestrin, C. XGBoost: A scalable tree boosting system. in Proceedings of the ACM SIGKDD International Conference on Knowledge Discovery and Data Mining vols 13-17-August-2016 785--794 (Association for Computing Machinery, 2016).29. Bengfort, B. et al. Yellowbrick v1.3. (2021) doi:10.5281/ZENODO.4525724.30. Eisenberg, D., Weiss, R. M., Terwilliger, T. C. \& Wilcox, W. Hydrophobic moments and protein structure. in Faraday Symposia of the Chemical Society vol. 17 109--120 (The Royal Society of Chemistry, 1982).31. Lee, B. \& Richards, F. M. The interpretation of protein structures: Estimation of static accessibility. J. Mol. Biol. 55, 379-IN4 (1971).32. Porter, C. T. \& Martin, A. C. R. BiopLib and BiopTools - A C programming library and toolset for manipulating protein structure. Bioinformatics 31, 4017--4019 (2015).33. Van Der Spoel, D. et al. GROMACS: Fast, flexible, and free. J. Comput. Chem. 26, 1701--1718 (2005).34. Nanni, L. \& Lumini, A. A new encoding technique for peptide classification. Expert Syst. Appl. 38, 3185--3191 (2011).35. Wong, T. T. Performance evaluation of classification algorithms by k-fold and leave-one-out cross validation. Pattern Recognit. 48, 2839--2846 (2015).36. Krstajic, D., Buturovic, L. J., Leahy, D. E. \& Thomas, S. Cross-validation pitfalls when selecting and assessing regression and classification models. J. Cheminform. 6, 10 (2014).37. Kohavi, R. A Study of Cross-Validation and Bootstrap for Accuracy Estimation and Model Selection. http://robotics.stanford.edu/\textasciitilde ronnyk (1995).38. Structure, B. M.-B. et B. A. (BBA)-P. \& 1975, undefined. Comparison of the predicted and observed secondary structure of T4 phage lysozyme. Elsevier.39. Chicco, D. Ten quick tips for machine learning in computational biology. BioData Mining vol. 10 35 (2017).40. Morris, J. The Precise Effect of Multicollinearity on Classification Prediction Ridge Regression View project Curriculum-Based VAM View project. https://www.researchgate.net/publication/336561703 (2014).41. Handling Multi-Collinearity in ML Models \textbar{} by Vishwa Pardeshi \textbar{} Towards Data Science. https://towardsdatascience.com/handling-multi-collinearity-6579eb99fd81.42. Jin Lee, C., Park, C.-S., Seok Kim, J. \& Baek, J.-G. A Study on Improving Classification Performance for Manufacturing Process Data with Multicollinearity and Imbalanced Distribution. 41, 25--33 (2015).43. Box, P. O., Van Der Maaten, L., Postma, E. \& Van Den Herik, J. Tilburg centre for Creative Computing Dimensionality Reduction: A Comparative Review Dimensionality Reduction: A Comparative Review. http://www.uvt.nl/ticc (2009).44. Vlachos, M. et al. Non-Linear Dimensionality Reduction Techniques for Classification and Visualization. (2002).45. Sorzano, C. O. S., Vargas, J. \& Pascual-Montano, A. A survey of dimensionality reduction techniques.46. Rifai, S., Vincent, P., Muller, X., Glorot, X. \& Bengio, Y. Contractive Auto-Encoders: Explicit Invariance During Feature Extraction. (2011).47. Chicco, D. \& Jurman, G. The advantages of the Matthews correlation coefficient (MCC) over F1 score and accuracy in binary classification evaluation. BMC Genomics 21, 6 (2020).48. Brown, J. B. Classifiers and their Metrics Quantified. Mol. Inform. 37, 1700127 (2018).49. Delgado, R. \& Tibau, X. A. Why Cohen's Kappa should be avoided as performance measure in classification. PLoS One 14, e0222916 (2019).50. Gorodkin, J. Comparing two K-category assignments by a K-category correlation coefficient. Comput. Biol. Chem. 28, 367--374 (2004).51. Wong, W. K., Leem, J. \& Deane, C. M. Comparative analysis of the CDR loops of antigen receptors Protein structure prediction View project Comparative analysis of the CDR loops of antigen. doi:10.1101/709840.}}\label{charles-a-janeway-j.-travers-p.-walport-m.-shlomchik-m.-j.-the-interaction-of-the-antibody-molecule-with-specific-antigen.-immunobiology-the-immune-system-in-health-and-disease.-garland-science-2001.2.-lu-r.-m.-et-al.-development-of-therapeutic-antibodies-for-the-treatment-of-diseases.-journal-of-biomedical-science-vol.-27-130-2020.3.-antibodies-market-size-share-trends-growth-forecast-2020-to-2025.-httpswww.marketdataforecast.commarket-reportsantibodies-market.4.-urquhart-l.-top-companies-and-drugs-by-sales-in-2020.-nat.-rev.-drug-discov.-2021-doi10.1038d41573-021-00050-6.5.-abhinandan-k.-r.-martin-a.-c.-r.-analyzing-the-degree-of-humanness-of-antibody-sequences.-j.-mol.-biol.-369-852862-2007.6.-north-b.-lehmann-a.-dunbrack-r.-l.-a-new-clustering-of-antibody-cdr-loop-conformations.-j.-mol.-biol.-406-228256-2011.7.-weitzner-b.-d.-dunbrack-r.-l.-gray-j.-j.-the-origin-of-cdr-h3-structural-diversity.-structure-23-302311-2015.8.-finn-j.-a.-et-al.-improving-loop-modeling-of-the-antibody-complementarity-determining-region-3-using-knowledge-based-restraints.-plos-one-11-e0154811-2016.9.-regep-c.-georges-g.-shi-j.-popovic-b.-deane-c.-m.-the-h3-loop-of-antibodies-shows-unique-structural-characteristics.-proteins-struct.-funct.-bioinforma.-85-13111318-2017.10.-maccallum-r.-m.-martin-a.-c.-r.-thornton-j.-m.-antibody-antigen-interactions-contact-analysis-and-binding-site-topography.-j.-mol.-biol.-262-732745-1996.11.-xu-j.-l.-davis-m.-m.-diversity-in-the-cdr3-region-of-vh-is-sufficient-for-most-antibody-specificities.-immunity-13-3745-2000.12.-kabat-e.-wu-t.-te-perry-h.-foeller-c.-gottesman-k.-sequences-of-proteins-of-immunological-interest.-1992.13.-almagro-j.-c.-et-al.-second-antibody-modeling-assessment-ama-ii.-proteins-structure-function-and-bioinformatics-vol.-82-15531562-2014.14.-almagro-j.-c.-et-al.-antibody-modeling-assessment.-proteins-struct.-funct.-bioinforma.-79-30503066-2011.15.-sircar-a.-kim-e.-t.-gray-j.-j.-rosettaantibody-antibody-variable-region-homology-modeling-server.-nucleic-acids-res.-37-w474w479-2009.16.-sivasubramanian-a.-sircar-a.-chaudhury-s.-gray-j.-j.-toward-high-resolution-homology-modeling-of-antibody-f-v-regions-and-application-to-antibody-antigen-docking.-proteins-struct.-funct.-bioinforma.-74-497514-2009.17.-leem-j.-dunbar-j.-georges-g.-shi-j.-deane-c.-m.-abodybuilder-automated-antibody-structure-prediction-with-datadriven-accuracy-estimation.-mabs-8-12591268-2016.18.-lepore-r.-olimpieri-p.-p.-messih-m.-a.-tramontano-a.-pigspro-prediction-of-immunoglobulin-structures-v2.-nucleic-acids-res.-45-w17w23-2017.19.-schoeder-c.-t.-et-al.-modeling-immunity-with-rosetta-methods-for-antibody-and-antigen-design.-60.20.-weitzner-b.-d.-gray-j.-j.-accurate-structure-prediction-of-cdr-h3-loops-enabled-by-a-novel-structure-based-c-terminal-constraint.-j.-immunol.-198-505515-2017.21.-choi-y.-deane-c.-m.-fread-revisited-accurate-loop-structure-prediction-using-a-database-search-algorithm.-proteins-struct.-funct.-bioinforma.-78-14311440-2010.22.-martin-a.-c.-r.-thornton-j.-m.-structural-families-in-loops-of-homologous-proteins-automatic-classification-modelling-and-application-to-antibodies.-j.-mol.-biol.-263-800815-1996.23.-almagro-j.-c.-et-al.-antibody-modeling-assessment.-proteins-struct.-funct.-bioinforma.-79-30503066-2011.24.-ferdous-s.-martin-a.-c.-r.-abdb-antibody-structure-databasea-database-of-pdb-derived-antibody-structures.-database-2018-2018.25.-abhinandan-k.-r.-martin-a.-c.-r.-analysis-and-prediction-of-vhvl-packing-in-antibodies.-protein-eng.-des.-sel.-23-689697-2010.26.-thomsen-m.-c.-f.-nielsen-m.-seq2logo-a-method-for-construction-and-visualization-of-amino-acid-binding-motifs-and-sequence-profiles-including-sequence-weighting-pseudo-counts-and-two-sided-representation-of-amino-acid-enrichment-and-depletion.-nucleic-acids-res.-40-2012.27.-shaner-m.-c.-blairs-i.-m.-schneider-t.-d.-sequence-logos-a-powerful-yet-simple-tool.28.-chen-t.-guestrin-c.-xgboost-a-scalable-tree-boosting-system.-in-proceedings-of-the-acm-sigkdd-international-conference-on-knowledge-discovery-and-data-mining-vols-13-17-august-2016-785794-association-for-computing-machinery-2016.29.-bengfort-b.-et-al.-yellowbrick-v1.3.-2021-doi10.5281zenodo.4525724.30.-eisenberg-d.-weiss-r.-m.-terwilliger-t.-c.-wilcox-w.-hydrophobic-moments-and-protein-structure.-in-faraday-symposia-of-the-chemical-society-vol.-17-109120-the-royal-society-of-chemistry-1982.31.-lee-b.-richards-f.-m.-the-interpretation-of-protein-structures-estimation-of-static-accessibility.-j.-mol.-biol.-55-379-in4-1971.32.-porter-c.-t.-martin-a.-c.-r.-bioplib-and-bioptools---a-c-programming-library-and-toolset-for-manipulating-protein-structure.-bioinformatics-31-40174019-2015.33.-van-der-spoel-d.-et-al.-gromacs-fast-flexible-and-free.-j.-comput.-chem.-26-17011718-2005.34.-nanni-l.-lumini-a.-a-new-encoding-technique-for-peptide-classification.-expert-syst.-appl.-38-31853191-2011.35.-wong-t.-t.-performance-evaluation-of-classification-algorithms-by-k-fold-and-leave-one-out-cross-validation.-pattern-recognit.-48-28392846-2015.36.-krstajic-d.-buturovic-l.-j.-leahy-d.-e.-thomas-s.-cross-validation-pitfalls-when-selecting-and-assessing-regression-and-classification-models.-j.-cheminform.-6-10-2014.37.-kohavi-r.-a-study-of-cross-validation-and-bootstrap-for-accuracy-estimation-and-model-selection.-httprobotics.stanford.eduronnyk-1995.38.-structure-b.-m.-b.-et-b.-a.-bba-p.-1975-undefined.-comparison-of-the-predicted-and-observed-secondary-structure-of-t4-phage-lysozyme.-elsevier.39.-chicco-d.-ten-quick-tips-for-machine-learning-in-computational-biology.-biodata-mining-vol.-10-35-2017.40.-morris-j.-the-precise-effect-of-multicollinearity-on-classification-prediction-ridge-regression-view-project-curriculum-based-vam-view-project.-httpswww.researchgate.netpublication336561703-2014.41.-handling-multi-collinearity-in-ml-models-by-vishwa-pardeshi-towards-data-science.-httpstowardsdatascience.comhandling-multi-collinearity-6579eb99fd81.42.-jin-lee-c.-park-c.-s.-seok-kim-j.-baek-j.-g.-a-study-on-improving-classification-performance-for-manufacturing-process-data-with-multicollinearity-and-imbalanced-distribution.-41-2533-2015.43.-box-p.-o.-van-der-maaten-l.-postma-e.-van-den-herik-j.-tilburg-centre-for-creative-computing-dimensionality-reduction-a-comparative-review-dimensionality-reduction-a-comparative-review.-httpwww.uvt.nlticc-2009.44.-vlachos-m.-et-al.-non-linear-dimensionality-reduction-techniques-for-classification-and-visualization.-2002.45.-sorzano-c.-o.-s.-vargas-j.-pascual-montano-a.-a-survey-of-dimensionality-reduction-techniques.46.-rifai-s.-vincent-p.-muller-x.-glorot-x.-bengio-y.-contractive-auto-encoders-explicit-invariance-during-feature-extraction.-2011.47.-chicco-d.-jurman-g.-the-advantages-of-the-matthews-correlation-coefficient-mcc-over-f1-score-and-accuracy-in-binary-classification-evaluation.-bmc-genomics-21-6-2020.48.-brown-j.-b.-classifiers-and-their-metrics-quantified.-mol.-inform.-37-1700127-2018.49.-delgado-r.-tibau-x.-a.-why-cohens-kappa-should-be-avoided-as-performance-measure-in-classification.-plos-one-14-e0222916-2019.50.-gorodkin-j.-comparing-two-k-category-assignments-by-a-k-category-correlation-coefficient.-comput.-biol.-chem.-28-367374-2004.51.-wong-w.-k.-leem-j.-deane-c.-m.-comparative-analysis-of-the-cdr-loops-of-antigen-receptors-protein-structure-prediction-view-project-comparative-analysis-of-the-cdr-loops-of-antigen.-doi10.1101709840.}}

\textbf{APPENDIX}

\end{document}